# Basics R

## Simple statistics

```{r, include=FALSE}
variable <- 1:10
value <- 5.50
```

```{r include=FALSE}
x <- 5
y <- 10
n <- 30
```

**Sample mean, standard deviation**

```{r}
mean <- mean(variable)
sd <- sd(variable)
```

**Removes values NA in a data set:**

```{r}
mean <- mean(variable, na.rm = TRUE)
sd <- sd(variable, na.rm = TRUE)
```

**Weighted mean & standard deviation**

```{r message=FALSE, warning=FALSE}
library(Hmisc)
weightedmean <- wtd.mean(x,y)
weightedsd <- sqrt(wtd.var(x,y))/sqrt(n)
```

**Variance**

```{r}
var(variable)
```

## Tables, frames & matrices

> As matrix

```{r message=FALSE, warning=FALSE}
matrix(c(1:8), nrow = 4, byrow = TRUE) #organized by row
matrix(c(1:8), ncol = 4, byrow = FALSE) #organized by col
```

> As data frame

```{r}
data.frame(Column1 = c(1:5), Column2 = c(1:5))
```

> As data table

```{r}
library(data.table)
data.table(Column1 = c(1:5), Column2 = c(1:5)) # or
data.table(matrix(c(1:8), nrow = 4)) # or
```

Transforming tables to other formats =

```{r}
table1 <- data.table(matrix(c(1:8), nrow = 4))
as.data.frame(table1)

table2 <- data.frame(Column1 = c(1:5), Column2 = c(1:5))
as.data.table(table2)
```

### Combining sets and setting names:

Beginning:

```{r}
table <- data.table(Column1 = 0:5, Column2 = 0:5)
newC <- 0:5
```

**Binding:**

```{r}
cbind(table, newC)
```

```{r}
rownames(table1) <- c("One", "Two", "Three", "Four")
colnames(table2) <- c("One", "Two")
```

Changing the order of a frequency table and factor

```{r}
value <- c("one", "three", "five", "one", "two", "three", "four", "two", "three")
table <- as.data.table(table(value))

table

table[,`value`:= factor(
 `value`, 
  levels = c(
    "one",
    "two",
    "three",
    "four",
    "five"
  )
)]

setorder(table, `value`)

table
```

**Frequencies & Probabilities**

```{r}
values <- c(1:5)
table(values)
prop.table(table(values))
round(prop.table(table(values)) * 100, 2)
```

## Data sets

Displaying head or tail of a data set:

```{r}
head(cars) # first 6 rows of the data set
tail(cars) # last 6 rows of the data set
```

Reading excel

`library(readxl)` / `data <- read.xls("data.xlsx", stringsAsFactors = TRUE)`

```{r}
names(cars) # shows the column names of the data set
attach(cars) # saves the names to be used as variables
```

### Subsetting

```{r}
cars[,1] # subsets by columns
cars[1,] # subsets by row
cars[cars$speed == 4,] # subsets by specific value 
cars[cars$speed != 4,] # subsets excluding specific value 
cars[cars$speed > 4,] # subsets over specific value 
cars[cars$speed < 5,] # subsets under specific value 
```

## R Markdown

```{r}
# Putting words in bold: **Word**
```

**Result**

```{r}
# Putting words in italic: *Word*
```

*Result*

```{r}
# dashes like this `here`
```

Show up like this: `here`

```{r}
# dashes like this with the letter r:  `r 4+4`
```

Asks r to have inline code. We can see the results here: `r 4+4`.

```{r}
# \newpage
```

Will start a new page for example in a pdf document

```{r}
# > 
# In here we can put a quote 
# >
```

> In here we can put a quote

### Setup rmarkdown & code chunks

+--------------------------+----------------------------------------------------------+
| Call                     | Description                                              |
+==========================+==========================================================+
| Warning = T/F \| Includ  | e / exclude warnings                                     |
+--------------------------+----------------------------------------------------------+
| Echo = T/F \| Includ     | e / exclude r chunks but show output                     |
+--------------------------+----------------------------------------------------------+
| Include = T/F \| Run cod | e but do not include in the knitted document             |
+--------------------------+----------------------------------------------------------+
| Comment = "" \| Include  | / exclude \#\# in output code chunks                     |
+--------------------------+----------------------------------------------------------+
| Message = T/F \| Includ  | es / excludes message from code                          |
+--------------------------+----------------------------------------------------------+
| out.width='100%'         | Adjusts size of figure / chart                           |
+--------------------------+----------------------------------------------------------+
| fig.width =              | Set specific size of figure / chart width                |
+--------------------------+----------------------------------------------------------+
| fig.height =             | Set specific size of figure / chart height               |
+--------------------------+----------------------------------------------------------+
| fig.cap=                 | Adds a title to a figure                                 |
+--------------------------+----------------------------------------------------------+
| fig.align=               | 'center', 'left', 'right', adjust figure / chart at page |
+--------------------------+----------------------------------------------------------+

### Latex

+-------------------+--------------------+-------------------------------+
| Symbol            | Writing            | Description                   |
+:=================:+:==================:+:=============================:+
| $\mu$             | `$\mu$`            | Population mean               |
+-------------------+--------------------+-------------------------------+
| $\sigma$          | `$\sigma$`         | Population standard deviation |
+-------------------+--------------------+-------------------------------+
| $\bar{x}$         | `$\bar{x}$`        | Sample mean                   |
+-------------------+--------------------+-------------------------------+
| ${e}$             | `${e}$`            | Standard error                |
+-------------------+--------------------+-------------------------------+
| $lambda$          | `$lambda$`         | Lambda                        |
+-------------------+--------------------+-------------------------------+
| $\ge$             | `$\ge$`            | Bigger than                   |
+-------------------+--------------------+-------------------------------+
| $\le$             | `$\le$`            | Smaller than                  |
+-------------------+--------------------+-------------------------------+
| $\pi$             | `$\pi$`            | Pie                           |
+-------------------+--------------------+-------------------------------+
| $\hat f$          | `$\hat f$`         | Estimator                     |
+-------------------+--------------------+-------------------------------+
| $\frac{1}{2}$     | `$\frac{1}{2}$`    | Fraction                      |
+-------------------+--------------------+-------------------------------+
| $\sum$            | `$\sum$`           | Sum                           |
+-------------------+--------------------+-------------------------------+
| $\sum_{i=1}^{n}$  | `$\sum_{i=1}^{n}$` | Sequence                      |
+-------------------+--------------------+-------------------------------+
| $\nabla$          | `$\nabla$`         | Gradient, Nabla               |
+-------------------+--------------------+-------------------------------+
| $\partial$        | `$\partial$`       | Partial                       |
+-------------------+--------------------+-------------------------------+
| $\beta$           | `$\beta$`          | Beta                          |
+-------------------+--------------------+-------------------------------+
| $\beta^1$         | `$\beta^1$`        | Beta 1                        |
+-------------------+--------------------+-------------------------------+
| $\alpha$          | `$\alpha$`         | Alpha                         |
+-------------------+--------------------+-------------------------------+
| $\approx$         | `$\approx$`        | Similar to                    |
+-------------------+--------------------+-------------------------------+

## Miscellaneous

```{r}
round(0.50, 2) # rounds a value to one decimals
rep(5,5) #repeats the number 5, 5 times
describe(variable) # basic information
fivenum(variable) # gives interquartile range & range
summary(variable) # basic statistics
str(variable) # describing the variable
dim(cars) # amount of rows and amount of columns
```

### Removing infinite + NA values

1.  Removing infinite values
2.  Removing NA values
3.  Changing infinite values to NA values

```{r}
variable[is.finite(variable)]
variable[is.na(variable)]
variable[is.infinite(variable)] <- NA
```

### Transforming variable types

```{r}
as.numeric(value)
as.character(value)
as.factor(value)
```
