# Basics

```{r, include=FALSE}
variable <- 1:10
value <- 5.50
```

```{r include=FALSE}
x <- 5
y <- 10
n <- 30
```

**Mathematical values**

$\mu$ = `$\mu$` = Population mean\
$\sigma$ = `$\sigma$` = Population sd\
$\bar{x}$ = `$\bar{x}` = Sample mean\
${e}$ = `${e}$` = Standard error\
$\pi$ = `$\pi$` = Pie $\ge$ = `$\ge$` = Bigger than\
$\le$ = `$\le$` = Smaller than

**Sample mean, standard deviation**

```{r}
mean <- mean(variable)
sd <- sd(variable)
```

Removes values NA in a data set:

```{r}
mean <- mean(variable, na.rm = TRUE)
sd <- sd(variable, na.rm = TRUE)
```

**Weighted mean & standard deviation**

```{r}
library(Hmisc)
weightedmean <- wtd.mean(x,y)
weightedsd <- sqrt(wtd.var(x,y))/sqrt(n)
```

**Variance**

```{r}
var(variable)
```

### Tables, frames & Matrices\*\*

As matrix =

```{r}
library(data.table)
matrix(c(1:8), nrow = 4, byrow = TRUE) #organized by row
matrix(c(1:8), ncol = 4, byrow = FALSE) #organized by col
```

As data frame =

```{r}
data.frame(Column1 = c(1:5), Column2 = c(1:5))
data.table(Column1 = c(1:5), Column2 = c(1:5))
```

As data table =

```{r}
data.table(matrix(c(1:8), nrow = 4)) # or
data.table(Variablex = 1:5, Variabley = 1:5)
```

Transforming tables from to other formats =

```{r}
table1 <- data.table(matrix(c(1:8), nrow = 4))
as.data.frame(table1)

table2 <- data.frame(Column1 = c(1:5), Column2 = c(1:5))
as.data.table(table2)
```

Binding and setting names =

rbind(table, newvariable) cbind(data, newvariable)

```{r}
rownames(table1) <- c("One", "Two", "Three", "Four")
colnames(table2) <- c("One", "Two")
```

```{r}
head(cars) # first 6 rows of the data set
tail(cars) # last 6 rows of the data set
```

### Removing infinite + NA values

1.  Removing Infinite values
2.  Removing NA values
3.  Changing Infinite values to NA values

```{r}
variable[is.finite(variable)]
variable[is.na(variable)]
variable[is.infinite(variable)] <- NA
```

### Transforming variable types

```{r}
as.numeric(value)
as.character(value)
as.factor(value)
```

**Read excel**

library(readxl) data \<- read.xls("data.xlsx", stringsAsFactors = TRUE)

### Setup rmarkdown:

+--------------------+-------------------------------------------------+
| Call               | Description                                     |
+====================+=================================================+
| Warning = FALSE    | Excluding warnings                              |
+--------------------+-------------------------------------------------+
| Echo=TRUE          | Include r chunks                                |
+--------------------+-------------------------------------------------+
|                    | Message = FALSE                                 |
+--------------------+-------------------------------------------------+
|                    | Fig.width =                                     |
+--------------------+-------------------------------------------------+
|                    | Fig.height =                                    |
+--------------------+-------------------------------------------------+
|                    | Fig.path =                                      |
+--------------------+-------------------------------------------------+
|                    | Error = TRUE                                    |
+--------------------+-------------------------------------------------+
|                    | Cache = FALSE                                   |
+--------------------+-------------------------------------------------+
|                    |                                                 |
+--------------------+-------------------------------------------------+
|                    |                                                 |
+--------------------+-------------------------------------------------+

### Miscellaneous:

```{r}
round(value, 2) # two decimals
rep(5,5) #repeats 5, 5 times
describe(variable)
fivenum(variable)
summary(variable)
str(variable) # describing the variable
dim(cars) # amount of rows and amount of columns
```
