# Basics

```{r, include=FALSE}
variable <- 1:10
value <- 5.50
```

```{r include=FALSE}
x <- 5
y <- 10
n <- 30
```

**Mathematical values**

$\mu$ = `$\mu$` = Population mean\
$\sigma$ = `$\sigma$` = Population sd\
$\bar{x}$ = `$\bar{x}` = Sample mean\
${e}$ = `${e}$` = Standard error\
$\pi$ = `$\pi$` = Pie $\ge$ = `$\ge$` = Bigger than\
$\le$ = `$\le$` = Smaller than

**Sample mean, standard deviation**

```{r}
mean <- mean(variable)
sd <- sd(variable)
```

Removes values NA in a data set:

```{r}
mean <- mean(variable, na.rm = TRUE)
sd <- sd(variable, na.rm = TRUE)
```

**Weighted mean & standard deviation**

```{r message=FALSE, warning=FALSE}
library(Hmisc)
weightedmean <- wtd.mean(x,y)
weightedsd <- sqrt(wtd.var(x,y))/sqrt(n)
```

**Variance**

```{r}
var(variable)
```

### Tables, frames & Matrices

As matrix =

```{r message=FALSE, warning=FALSE}
library(data.table)
matrix(c(1:8), nrow = 4, byrow = TRUE) #organized by row
matrix(c(1:8), ncol = 4, byrow = FALSE) #organized by col
```

As data frame =

```{r}
data.frame(Column1 = c(1:5), Column2 = c(1:5))
data.table(Column1 = c(1:5), Column2 = c(1:5))
```

As data table =

```{r}
data.table(matrix(c(1:8), nrow = 4)) # or
data.table(Variablex = 1:5, Variabley = 1:5)
```

Transforming tables from to other formats =

```{r}
table1 <- data.table(matrix(c(1:8), nrow = 4))
as.data.frame(table1)

table2 <- data.frame(Column1 = c(1:5), Column2 = c(1:5))
as.data.table(table2)
```

Binding and setting names =

rbind(table, newvariable) cbind(data, newvariable)

```{r}
rownames(table1) <- c("One", "Two", "Three", "Four")
colnames(table2) <- c("One", "Two")
```

```{r}
head(cars) # first 6 rows of the data set
tail(cars) # last 6 rows of the data set
```

Changing the order of a frequency table and factor

```{r}
value <- c("one", "three", "five", "one", "two", "three", "four", "two", "three")
table <- as.data.table(table(value))

table

table[,`value`:= factor(
 `value`, 
  levels = c(
    "one",
    "two",
    "three",
    "four",
    "five"
  )
)]

setorder(table, `value`)

table
```


### Removing infinite + NA values

1.  Removing Infinite values
2.  Removing NA values
3.  Changing Infinite values to NA values

```{r}
variable[is.finite(variable)]
variable[is.na(variable)]
variable[is.infinite(variable)] <- NA
```

### Transforming variable types

```{r}
as.numeric(value)
as.character(value)
as.factor(value)
```

**Read excel**

library(readxl) data \<- read.xls("data.xlsx", stringsAsFactors = TRUE)

### Markdown

```{r}
# Putting words in bold: **Word**
```

**Result**

```{r}
# Putting words in italic: *Word*
```

*Result*

```{r}
# dashes like this `here`
```

Show up like this: `here`

```{r}
# dashes like this with the letter r:  `r 4+4`
```

Asks r to have inline code. We can see the results here: `r 4+4`.

```{r}
# \newpage
```

Will start a new page for example in a pdf document

```{r}
# > 
# In here we can put a quote 
# >
```

>
In here we can put a quote
>


### Setup rmarkdown & code chunks

+----------------------+----------------------------------------------------------+
| Call                 | Description                                              |
+======================+==========================================================+
| Warning = TRUE/FALSE | Include / exclude warnings                               |
+----------------------+----------------------------------------------------------+
| Echo = TRUE/FALSE    | Include / exclude r chunks but show output               |
+----------------------+----------------------------------------------------------+
| Include = TRUE/FALSE | Run code but do not include in the knitted document      |
+----------------------+----------------------------------------------------------+
| Comment = "" | Include / exclude \#\# in output code chunks             |
+----------------------+----------------------------------------------------------+
| Message = TRUE/FALSE | Includes / excludes message from code                    |
+----------------------+----------------------------------------------------------+
| out.width='100%'     | Adjusts size of figure / chart                           |
+----------------------+----------------------------------------------------------+
| fig.width =          | Set specific size of figure / chart width                |
+----------------------+----------------------------------------------------------+
| fig.height =         | Set specific size of figure / chart height               |
+----------------------+----------------------------------------------------------+
| fig.cap=             | Adds a title to a figure                                 |
+----------------------+----------------------------------------------------------+
| fig.align=           | 'center', 'left', 'right', adjust figure / chart at page |
+----------------------+----------------------------------------------------------+


### Miscellaneous

```{r}
round(0.50, 2) # rounds a value with two decimals
rep(5,5) #repeats the number 5, 5 times
describe(variable)
fivenum(variable)
summary(variable)
str(variable) # describing the variable
dim(cars) # amount of rows and amount of columns
```


### Subsetting

```{r}
cars[,1] # subsets by columns
cars[1,] # subsets by row
cars[cars$speed == 4,] # subsets by specific value 
cars[cars$speed > 4,] # subsets over specific value 
cars[cars$speed < 5,] # subsets under specific value 
```

