# Basics

```{r}
variable <- 1:10
```

```{r include=FALSE}
x <- 5
y <- 10
n <- 30
```

**Mathmatical values**

$\mu$ = Population mean   
$\sigma$ = Population sd   
$\bar{x}$ = Sample mean   
${e}$ = Standard error 
$\pi$ = Pie 
$\ge$ = Bigger than     
$\le$ = Smaller than 

**Sample mean, standard deviation**

```{r}
mean(variable)
sd(variable)
```

Removes values NA in a data set:

```{r}
mean(variable, na.rm = TRUE)
sd(variable, na.rm = TRUE)
```

**Weighted mean & standard deviation**

```{r}
library(Hmisc)
weightedmean <- wtd.mean(x,y)
weightedsd <- sqrt(wtd.var(x,y))/sqrt(n)
```

**Variance**

```{r}
var(variable)
```

**Tables frames & Matrixes**

As matrix =

```{r}
library(data.table)
matrix(c(1:8), nrow = 4, byrow = TRUE) #organized by row
matrix(c(1:8), ncol = 4, byrow = FALSE) #organized by col
```

As data frame = 

```{r}
data.frame(Column1 = c(1:5), Column2 = c(1:5))
data.table(Column1 = c(1:5), Column2 = c(1:5))
```

As data table =

```{r}
data.table(matrix(c(1:8), nrow = 4)) # or
data.table(Variablex = 1:5, Variabley = 1:5)
```

Transforming tables from to other formats =

```{r}
table1 <- data.table(matrix(c(1:8), nrow = 4))
as.data.frame(table1)

table2 <- data.frame(Column1 = c(1:5), Column2 = c(1:5))
as.data.table(table2)
```

Binding and setting names = 

rbind(table, newvariable)
cbind(data, newvariable)

```{r}
rownames(table1) <- c("One", "Two", "Three", "Four")
colnames(table2) <- c("One", "Two")
```

**Other:**

```{r}
value <- 5.50
round(value, 2) # two decimals
as.numeric(value)
rep(5,5) #repeats 5, 5 times
describe(variable)
fivenum(variable)
summary(variable)
str(variable) #explains the variable
```

**Read excel**

library(readxl)
data <- read.xls("data.xlsx", stringsAsFactors = TRUE)

