# Simple regressions

```{r include=FALSE}
data <- cars
x <- cars$speed
y <- cars$dist
d <- rep(0,50)
```

## Basics regressions

Regressions, correlation and dummy's

Y = Dependent  
X = Explanatory

**Correlation**

```{r}
cor(x, y)
```

**Creating the regression:**

1. To plot the regression model
2. Evaluates the coefficient of the model
3. Only the first colum estimattion 

```{r}
model <- lm(y~x, data = data)
summary(model)$coef
est <- summary(model)$coef[,1]
```


### Summarizing regressions:

1. Using stargazer package

```{r message=FALSE, warning=FALSE}
library(stargazer)

stargazer(lm(y~x, data=data), type="text")

# Multiple models adjacent

model1 <- lm(y~x, data=data)
model2 <- lm(x~y, data=data)

stargazer(model1, model2, type="text")

```

2. Using summary function:

```{r}
summary(lm(y~x))
```

Regressions

**Plotting regression**

```{r}
plot(y~x,data=data, 
     main="Title",
     ylab="yname",
     xlab="xname"
     )
```

**Including regression line:**

```{r}
plot(y~x,data=data, 
     main="Title",
     ylab="yname",
     xlab="xname"
     )
abline(lm(y~x, data=data), col="blue")
```

**Confidence interval around slope**

```{r}
confint(lm(y~x), level=0.95)
```

**Sub-sampling regression**

Specify dimensions [,]. 
First is row. Column, second. 

1. Selects the rows where age is larger than 5.
2. Lower than 5.

```{r}
sub1 <- summary(lm(y~x, data=data["speed">=5,]))
sub2 <- summary(lm(y~x, data=data["speed"<=5,]))
```

###Dummy variables, diff in means


### Regression + dummy

Y = Constant0 + B0 * X - Diff in means + B1 * variable1*2


**Omitting the intercept:**

Shows the means separately and not the difference between means. Tests whether the expected counts are different from zero. 

```{r}
lm(y ~ x - 1, data = data)
```

Reorders group, to specific value to be first.
 
 variable2 <- relevel(variable, "C")

## Prediction  

``` {r}
model <- lm(y~x)

newdata <- data.frame(variablename = c(1:50))
pred <- predict(model, newdata = newdata)
```

**Prediction confidence interval:**

1. One value
2. Multiple values from a existing data frame

```{r}
pred1 <- predict(model, data.frame(valuename = x), interval = "confidence", level=0.95)
pred2 <- predict(model, newdata = newdata, interval = "confidence", level=0.95)
```

**Prediction interval**

1. One value
2. Multiple values from a existing data frame

```{r}
pred1 <- predict(model, data.frame(valuename = x), interval="predict",level=0.95)
pred2 <- predict(model, newdata, interval="predict",level=0.95)
```


### Confidence and prediction plotting

Adds: observed values, fitted line, conf interval, predicted interval

```{r message=FALSE, warning=FALSE}
library(HH)
fit <- lm(y~x, data = data)
ci.plot(fit)
```

###Prediction with dummy variables

Prediction = ð›¼1+ð›¼2Constant Dummy+ð›½1ð‘†ð‘–ð‘§ð‘’+ð›½2Slope Dummy

###Prediction intervals examples

**Prediction**

```{r}
fit <- lm(y ~ x + d + d, data = data)

pred <- predict(fit, data.frame(VariableX = c(10), Dummy = c(1), Slopedummy = c(10)) )
```
 
**Confidence interval prediction**  

```{r}
fit <- lm(y ~ x + d + d, data = data)

pred <- predict(fit, data.frame(VariableX = c(10), Dummy = c(1), Slopedummy = c(10), interval="confidence"))
```  

**Prediction interval**

```{r}
fit <- lm(y ~ x + d + d, data = data)

pred <- predict(fit, data.frame(VariableX = c(10), Dummy = c(1), Slopedummy = c(10), interval="predict"))
```


## Data problems

**Residual plot**

```{r}
# residual.plots(fitted(fit), resid(fit), sigma.hat(fit), main="Title")
```

**Influential measure test**

```{r}
im <- influence.measures(fit)
```


###Multicollinearity

1. F-test
2. Variance inflation factors greater than 10

```{r}
fit <- lm(y~x + d, data = data)

anova(fit)

vif(fit)
```

### ANOVA

**One-way:** one value

```{r}
res.aov <- aov(y ~ x, data = data)
summary(res.aov)
```

**Two-way:** more than two factors 

```{r}
res.aov <- aov(y ~ x + d, data = data)
summary(res.aov)
```

With interaction

```{r}
res.aov <- aov(y ~ x * d, data = data)
summary(res.aov)
```

**Three-way**

1. Three way
2. With interaction

```{r}
summary(aov(y ~ x + d, data=data))
summary(aov(y ~ x + d, data=data))
```

**MANOVA**: Multiple factors

1. Test in difference
2. Test separately


test_manova <- manova(cbind(y, d) ~ x, data = data)
summary(test_manova)
summary.aov(test_manova)

