---
editor_options: 
  markdown: 
    wrap: 72
---

# Basics Python

```{python, warning=FALSE, message=FALSE, comment=FALSE}
import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(reticulate)
os <- import("os")
os$listdir(".")
options(reticulate.repl.quiet = FALSE)
```

variable using instead of <-, equal sign =.

```{r}
y = 5 + 5
y
```

Printing characters

```{r}
print('Hello, readers!')
```

Printing numbers

```{r}
print(15)
```

Printing length of a value

```{python, warning=FALSE, message=FALSE}
Length = len('Danielle')
print(Length)
```


### Data set


Loading the data set & viewing head + tails:

```{python}
sns.set_context('paper')

tips = sns.load_dataset('tips')

tips.head()
tips.tail()
```

1. Length of the variable
2. Shape: Number of rows and number of columns
3. Type of variables + basic info
3. Descriptive statistics variable

```{python}
len(tips)

tips.shape

tips.info()

tips.describe()
```

### Matrixes

Series:

Panda series method:

```{python}
Serie1 = pd.Series(
  [4200, 8000, 6500],
  index=["Amsterdam", "Toronto", "Tokyo"]
  )
Serie1
```

Python dictionary method:

```{python}
Serie2 = pd.Series({"Amsterdam": 5, "Tokyo": 8})
Serie2
```


Data frame:

```{python}
Combined_serie = pd.DataFrame({
  "Revenue": Serie1,
  "Employee_count": Serie2
  })
  
Combined_serie
```

Sub-setting by row:

```{python}
Combined_serie["Tokyo":]
```


### Creating our own functions:

Saying hello + name

```{python, warning=FALSE, message=FALSE}
def printing_name(name):
  print('Good morning,', name)
printing_name('Danielle')
```

Multiple arguments:

Saying hello + name + location

```{python, warning=FALSE, message=FALSE}
def welcome(name, location):
  print("Good morning", name, "Welcome to", location)
welcome("Danielle,", "class.")
```

### Bar plot


```{python}
sns.set_context('paper')

tips = sns.load_dataset('tips')
tips.head()

sns.barplot(x = 'day', y = 'total_bill', hue = 'sex', data = tips,
            palette = 'Blues', edgecolor = 'w')
tips.groupby(['day','sex']).mean()

plt.show()
```

