---
title: "Summary and examples of r-code"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
site: bookdown::bookdown_site
documentclass: book
bibliography:
- book.bib
- packages.bib
biblio-style: apalike
link-citations: yes
description: This is a summary of r code learned throughout several courses of my
  master in management.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source = "watch-out", options(scipen=999), out.width = "60%", comment = FALSE) 
source("www/init.R")
source("www/Table_Design.R")
```

# Prerequisites

This is a *sample* book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
