---
title: "Summary and examples of r-code"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
site: bookdown::bookdown_site
documentclass: book
bibliography:
- book.bib
- packages.bib
biblio-style: apalike
link-citations: yes
description: This is a summary of r code learned throughout several courses of my
  master in management.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source = "watch-out", options(scipen=999), out.width = "100%", comment = "", warning=FALSE) 
source("www/init.R")
source("www/Table_Design.R")
```

# Preface {.unnumbered}

The following document has been prepared to have a prompt link to the code learned and used throughout several projects in the last year. As a beginner in programming, it is useful to have input accessible to avoid having to look through all repositories.

------------------------------------------------------------------------

The **bookdown** package that has been used for this format can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

To compile this example to PDF, you will need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].


*Prepared by: Daniëlle Kotter*
